### 2017-03-13 CO750-ApproximationAlgorithmsTopics/assignments/assignment_02.tex
improved notation, and thus clarity in 3b

We begin by working towards a compact formulation for $(IP)$ on $G$. Since $G$ is a cycle we may label the vertices of $G$ as $V(G) = \{v_1,\dots, v_n\}$ such that $v_iv_j \in E(G)$ if and only if $$j \equiv i\pm1\mod n$$. For $i, \ell \in [n]$ ($[n] := \{1,\dots, n\}$) we define the path from $i$ of length $\ell$ as 
$$P(i, \ell) := \{v_i, v_{i+1}, \dots, v_{i + \ell} \}$$
where addition is taken modulo $n$. We use the convention that when $\ell = n$ we have the entire cycle, and when $\ell = 0$ we have just $v_i$. We define the family of all paths on $G$ as
$$\cP := \{P(i,\ell) : i, \ell \in [n]\}.$$
Observe that the size of $\cP$ is polynomial in $n$, in particular $|\cP| = O(n^2)$ (since $n$ choices for $i$ and $n$ choices for $\ell$ covers all possible $P(i,\ell)$ with some overlap).

@@

Let $x$ be a feasible solution for $(\overline{IP})$. Let $S \subseteq V$. If we consider the connected components of $G[S]$ we observe that they are all paths. Hence there exist disjoint $P_1, \dots, P_k \in \cP$ such that

@@

>>>>

We begin by working towards a compact formulation for $(IP)$ on $G$. Since $G$ is a cycle we may label the vertices of $G$ as $V(G) = \{v_1,\dots, v_n\}$ such that $v_iv_j \in E(G)$ if and only if $$j \equiv i\pm1\mod n$$. For $v_i, v_j \in V(G)$ we define the path from $v_i$ to $v_j$ as 
$$P(i, \ell) := \{v_i, v_{i+1}, \dots, v_{j} \}$$
where addition is taken modulo $n$. We define the family of all paths on $G$ as
$$\cP := \{P(v_i, v_j) : v_i, v_j \in V(G)\}.$$
Observe that the size of $\cP$ is polynomial in $n$, in particular $|\cP| = O(n^2)$ (since there are $n$ choices for $v_i$ and $n$ choices for $v_j$).

@@

Let $x$ be a feasible solution for $(\overline{IP})$. Let $S \subseteq V$. If $S = V$ then $f(S)= f(V)=0\leq x(\delta(S)) $ (otherwise $(\overline{IP})$ and $(IP)$ are infeasible). By symmetry of $f$ the same holds for $S =\emptyset$. Thus we may assume $\emptyset \subset S \subset V$. If we consider the connected components of $G[S]$ we observe that they are all paths. Hence there exist disjoint $P_1, \dots, P_k \in \cP$ such that

@@

We make an observation about optimal solutions to $(\overline{IP})$ which contain a $0$-valued edge. This will lead us to an approximation algorithm with a better bound than the one found in $3(a)$.
\begin{lemma}\label{lemma:0edge}
Let $x$ be an optimal solution $(\bar{IP})$. Suppose there exists $v_iv_{i+1} \in E(G)$ such that $x(v_iv_{i+1}) = 0$. Reindex (by subtracting $i$ from each index) $V(G)$ so that $i = 0$. That is $x(v_0v_1)= 0$. Then for all $v_jv_{j+1} \in E(G)$, $j \neq 0$,
$$x(v_jv_{j+1}) = f(P(v_1, v_j)).$$
For each $v_jv_{j+1} \neq v_0v_{1} \in E(G)$. Since $x$ is feasible and $x(v_0v_{1}) = 0$,
$$f(P(v_1, v_j)) \leq x(\delta(P(v_1,v_j))) = x(v_jv_{j+1}) + x(v_0v_{1}) = x(v_jv_{j+1}).$$
Suppose for a contradiction there exists $e_j = v_{j}v_{j+1} \in E(G)$ such that
$$f(P(v_1,v_j)) < x(v_jv_{j+1}).$$
Since $f(V) = 0$, $j\neq 0$. Let $x' \in \Z^{|E|}$ we given by
$$x'(e) = \begin{cases}
x(e), &\text{ if $e\neq v_jv_{j+1}$} \\
f(P(v_1,v_{j})), &\text{if $e=v_jv_{j+1}$}. 
We claim that $x'$ is feasible for $(\overline{IP})$. By the feasibility of $x$, we need only check cut constraints involving $v_jv_{j+1}$. Let $v_i \in V(G)$, such that $i \neq j-1$. We consider $P(v_i,v_j)\in \cP$.
First consider the case where $0 < i \leq j$. In this case $v0,v_1 \in P(v_{j+1}, v_{i-1})$. We compute:
f(P(v_i,v_j)) &= f(P(v_{j+1}, v_{i-1})) &\text{by symmetry} \\
&\leq f(P(v_{j+1}, v_0)) + f(P(v_{1},v_{i-1})) &\text{by subadditivity} \\
&= f(P(v_1, v_j)) + f(P(v_1, v_{i-1})) &\text{by symmetry} \\
&= x'(v_jv_{j+1}) + f(P(v_1, v_{i-1})) &\text{by definition of $x'$} \\
&\leq x'(v_jv_{j+1}) + x(v_1v_{i-1}) +x(v_0v_1) &\text{by feasibility of $x$} \\
&= x'(v_jv_{j+1}) + x'(v_1v_{i-1}) &\text{by definition of $x'$, and $x(v_0v_1)=0$} \\
&=x'(\delta(P(v_i,v_j))).
Hence $x'$ satisfies the cut constraint.
Now consider the case where $i > j$. In this case $v_0,v_1 \in P(v_i,v_j)$. We compute (similarly):
f(P(v_i,v_j))&\leq f(P(v_{i}, v_0)) + f(P(v_{1},v_{j})) &\text{by subadditivity} \\
&= f(P(v_1, v_{i-1})) + f(P(v_1, v_{j})) &\text{by symmetry} \\
&= f(P(v_1,v_{i-1})) + x'(v_jv_{j+1}) &\text{by definition of $x'$} \\
&\leq x(v_1v_{i-1}) + x(v_0v_1) + x'(v_jv_{j+1}) &\text{by feasibility of $x$} \\
&= x'(v_jv_{j+1}) + x'(v_1v_{i-1}) &\text{by definition of $x'$, and $x(v_0v_1)=0$} \\
&=x'(\delta(P(v_i,v_j))).
Therefore in either case $x'$ is feasible. But
\sum_{e\in E} c(e) x(e) &= c(v_jv_{j+1})x(v_jv_{j+1}) + \sum_{e\in E\backslash\{v_jv_{j+1}\}} c(e)x(e) \\
&> c(v_jv_{j+1})x'(v_jv_{j+1}) +  \sum_{e\in E\backslash\{v_jv_{j+1}\}} c(e)x(e) &\text{since $x'(v_jv_{j+1}) = f(P(v_1,v_j) < x(v_jv_{j+1})$} \\
&= c(v_jv_{j+1})x'(v_jv_{j+1}) + \sum_{e\in E\backslash\{v_jv_{j+1}\}} c(e)x'(e) &\text{by our choice of $x'$} \\
&= \sum_{e\in E} c(e) x'(e),
contradicting the optimality of $x$.
This lemma tells us precisely how to construct optimal solutions to $(\overline{IP})$ in the presence of $0$-valued edges.
